Endogenous Politics and \\ the Design of Trade Institutions

### 2017-01-09 to-do.tex
add notes from Wouters & Zissimos (2016)
outline approach to introducing my obj fcn in end of sec 3

		\item In general it seems to me that you seem to ``dependent" on Bagwell and Staiger... (e.g. beginning of Sections 4 and 6).

@@

						\item Denote the government's welfare function as $W_G(\tau,\tau^*,e(\tau)) = W(\tau,\tau^*) + g(e(\tau))$ where $W = CS_X + PS_X + CS_Y + PS_Y +TR$ is social welfare. Then do analysis...
					\end{itemize}
			\end{itemize}
		\item Comparison to GH94
			\begin{itemize}
				\item Contingent payments are hard to reconcile with uncertainty
				\item It's the truthful contribution schedule in GH94 that implies the constant weighting of $1+a$ for sectors represented by a lobby. Won't get this clean, constant weighting with other schedules.
				\item Gov't doesn't care about \textit{anything} because it gets the same payoff no matter what
					\begin{itemize}
						\item With one lobby, this lobby-has-all-the-bargaining-power feature is an equilibrium implication of the menu-auction set-up where the lobby makes a TIOLI offer
						\item Gov't gets all the surplus when there are two or more lobbies
						\item This result holds in DGH97, p. 767. Footnote 10 explicitly says that this is because lobby makes TIOLI offer.
						\item So go to MRC07 where bargaining power can be manipulated
					\end{itemize}
				\item Menu auction eqm requirement: $(C(\tau),\tau)$ must maximize joint surplus
					\begin{itemize}
						\item NB: $\tau$ that maximizes gov't welfare doesn't generically maximize joint surplus (unless it also maximizes lobby payoff)
						\item In my model, optimum for lobby would be prohibitive tariff, $e=0$. Joint optimum determined by $\frac{\partial \ga}{\partial e}\pi_X = 2$ and $\frac{\partial CS_X}{\partial \tau} + \left[ 1 + \ga(e)\right]\frac{\partial \pi_X}{\partial \tau} + \frac{\partial TR}{\partial \tau} = 0$. Not clear how this is related to lobby's problem, but likely to have higher $\tau$ and lower $e$ than government's optimum (unless $\ga$ changes very steeply between the two points $e_J$ and $e_G$)
						\item Careful of transferable vs. non-transferable utility
					\end{itemize}
				\item Not sure if contingent vs. non-contingent payments are a big deal
				\item If I change nothing else but subtract $C$ from gov't welfare function, no qualitative change b/c of 0 bargaining weight
			\end{itemize}
		\item MRC07
			\begin{itemize}
				\item Need case of fixed capital ($z=0$)
				\item NB (p. 1391): optimal trade agreement if lobbies have all the bargaining power and no ex-ante lobbying is free trade
					\begin{itemize}
						\item Page 26-28 of working paper has government with all the bargaining power (and no capital mobility, but remember that the capital owners choose where to invest in the first period). Here, could be free trade but could be interior to get some rents.
					\end{itemize}
				\item Then re-do MRC07 with $W(\tau) + g(e)$, and version where I subtract $e$
					\begin{itemize}
						\item Should have gov't make lobby indifferent and maximize joint surplus. But again, this doesn't necessarily maximize gov't welfare
						\item Or rather, should I follow Limao and Tovar?
					\end{itemize}
			\end{itemize}
	\end{itemize}

@@

\section{Lit review}
\begin{itemize}
	\item Find MRC cites (maybe some in Ethier) for why PE shocks can't be addressed
		\begin{itemize}
			\item Bagwell and Staiger 1999 AER
		\end{itemize}
	\item Have to make connection to endogenous papers clearer
		\begin{itemize}
			\item and make clear that GH is isomorphic / provides microfoundations for Baldwin objective function
			\item More on Coates $\&$ Ludema (2001)
		\end{itemize}
	\item I think that I need to bring people along more carefully from GH/MRC to what I'm doing
	\item Paola: How do I want to sell the paper?
		\begin{itemize}
			\item In relation to what literature?
				\begin{itemize}
					\item What literature does endogenous? Coates-Ludema 2001, MRC, GH
					\item What literature \textit{doesn't}?
				\end{itemize}
			\end{itemize}
	\item Review Bown, Bagwell $\&$ Staiger
	\item Do I include papers like BS2001 (non-repeated) in lit review?
	\item Read Rosendorff $\&$ Milner (2001) [from Cristiane Carneiro at PEIO]
	\item Kyle: McLaren in his chapter calls the Grossman and Helpman setup with
	 contributions `felonious'
	\item David DeRemer: Safeguards have dropped off in use around the world
		\begin{itemize}
			\item Look at U.S. vs. ROW (he sent papers)
		\end{itemize}
	\item Limao and Tovar 2011
		\begin{enumerate}
			\item Their 2nd (and final) stage is Nash bargaining. Is there an isomorphism to my setup?
			\item It IS clear to take their first, limited case of commitment to tariff cap only (ignore NTB) as analogy for gov't in my model choosing commitment to different tariff level. Need to answer Q1 above in order to answer points 3 and 4:
			\item Their gains from commitment derive from improvement in bargaining power vis-a-vis the lobby (Schelling conjecture invoked). How does this relate to the gain from commitment in my model?
			\item Top of page 6: ``large enough bargaining power.'' Can I make connection to $\ga(e)$?
			\item Note: in their equation 3, $C$ enters expression for NBS's ($\tau$) in equilibrium, it doesn't disappear as in GH
		\end{enumerate}
\end{itemize}

@@

\section{Add empirics / evidence?}

>>>>

@@

		\item Comparison to GH94
				\item Contingent payments are hard to reconcile with uncertainty
				\item It's the truthful contribution schedule in GH94 that implies the constant weighting of $1+a$ for sectors represented by a lobby. Won't get this clean, constant weighting with other schedules.
				\item Gov't doesn't care about \textit{anything} because it gets the same payoff no matter what
						\item With one lobby, this lobby-has-all-the-bargaining-power feature is an equilibrium implication of the menu-auction set-up where the lobby makes a TIOLI offer
						\item Gov't gets all the surplus when there are two or more lobbies
						\item This result holds in DGH97, p. 767. Footnote 10 explicitly says that this is because lobby makes TIOLI offer.
						\item So go to MRC07 where bargaining power can be manipulated
				\item Menu auction eqm requirement: $(C(\tau),\tau)$ must maximize joint surplus
						\item NB: $\tau$ that maximizes gov't welfare doesn't generically maximize joint surplus (unless it also maximizes lobby payoff)
						\item In my model, optimum for lobby would be prohibitive tariff, $e=0$. Joint optimum determined by $\frac{\partial \ga}{\partial e}\pi_X = 2$ and $\frac{\partial CS_X}{\partial \tau} + \left[ 1 + \ga(e)\right]\frac{\partial \pi_X}{\partial \tau} + \frac{\partial TR}{\partial \tau} = 0$. Not clear how this is related to lobby's problem, but likely to have higher $\tau$ and lower $e$ than government's optimum (unless $\ga$ changes very steeply between the two points $e_J$ and $e_G$)
						\item Careful of transferable vs. non-transferable utility
				\item Not sure if contingent vs. non-contingent payments are a big deal
				\item If I change nothing else but subtract $C$ from gov't welfare function, no qualitative change b/c of 0 bargaining weight
		\item MRC07
				\item Need case of fixed capital ($z=0$)
				\item NB (p. 1391): optimal trade agreement if lobbies have all the bargaining power and no ex-ante lobbying is free trade
						\item Page 26-28 of working paper has government with all the bargaining power (and no capital mobility, but remember that the capital owners choose where to invest in the first period). Here, could be free trade but could be interior to get some rents.
				\item Then re-do MRC07 with $W(\tau) + g(e)$, and version where I subtract $e$
						\item Should have gov't make lobby indifferent and maximize joint surplus. But again, this doesn't necessarily maximize gov't welfare
						\item Or rather, should I follow Limao and Tovar?
		\item Approach: denote the government's welfare function as $W_G(\tau,\tau^*,e) = W(\tau,\tau^*) + g(e)$ where $W = CS_X + PS_X + CS_Y + PS_Y +TR$ is social welfare.
				\item This is like DGH97 (they write as $G(a,C)$, so no problem putting an $e$ in there), BS1999 `Economic Theory of GATT' (I think), $W(p,\tilde{p}^w)$
						\item I've already written $W(\ga(e,s),\tau,\tau^*)$
						\item I think people are most concerned with underlying economic model. I need a numeraire anyway. DGH97 doesn't have numeraire b/c small country
				\item Then specialize to BS2005 when necessary (tariff cap example, picture for interior max of gov't obj fcn)
				\item Menu auction framework doesn't work with one lobby because gov't doesn't get any surplus. Switch to MRC07 (bargaining). But want decreasing returns (LT).
				\item Make argument that want this version like Baldwin-style literature for comparison
						\item Want to make argument that things go wrong in standard model when you insert $e$ into $\ga$ (also, when $e$ goes into $\ga$ need to subtract $e$ too, and this often results in interior optimum)
						\item This is an easier way to get the result from MRC07 / LT, can make comparisons, develop design results more easily
				\item Want non-contingent set-up for different assumptions on timing / meaning of $s$; contingent only makes sense if $e$ chosen after $s$ and $s$ is completely verifiable to lobby. If $s$ is political uncertainty, it probably realizes after/contemporaneously with $e$
						\item I can leave general, then specify for escape clause. Or I can just go with my order of $s$ then $e$, which assumes $s$ is an economic shock.
				\item Bargaining framework is not appealing when gov't is non-unitary
				\item Limao $\&$ Tovar: lobby objective the same as mine. $G = \Psi(C) + aW(p)$. Gov't decides whether to commit in int'l agreement, then Nash bargains with lobby over $C$ and policies $(t,\tau)$ where $\tau$ is a NTB
						\item Section 3.3: They show exactly what I need to show. If gov't can commit with this obj fcn, it will as long as it doesn't have all the bargaining power and there are decreasing returns to lobbying.
						\item Gov't does not have an interest in maximizing joint surplus, because it doesn't appropriate all the gains
				\item Only difference between Limao / Tovar and mine is $\Psi(e)$ vs. $[\ga(e) - 1]\pi_x(\tau)$
						\item This means there's a positive cross partial between $e$ and $\tau$ in my formulation, so that an incraese in $\tau$ increases $W$ more the higher is $e$

@@

\section{Lit review}
\begin{itemize}
	\item Find MRC cites (maybe some in Ethier) for why PE shocks can't be addressed
		\begin{itemize}
			\item Bagwell and Staiger 1999 AER
		\end{itemize}
	\item Have to make connection to endogenous papers clearer
		\begin{itemize}
			\item and make clear that GH is isomorphic / provides microfoundations for Baldwin objective function
			\item More on Coates $\&$ Ludema (2001)
			\item Check out Ben and James new paper
		\end{itemize}
	\item I think that I need to bring people along more carefully from GH/MRC to what I'm doing
	\item Paola: How do I want to sell the paper?
		\begin{itemize}
			\item In relation to what literature?
				\begin{itemize}
					\item What literature does endogenous? Coates-Ludema 2001, MRC, GH
					\item What literature \textit{doesn't}?
				\end{itemize}
			\end{itemize}
	\item Review Bown, Bagwell $\&$ Staiger
	\item Do I include papers like BS2001 (non-repeated) in lit review?
	\item Read Rosendorff $\&$ Milner (2001) [from Cristiane Carneiro at PEIO]
	\item Kyle: McLaren in his chapter calls the Grossman and Helpman setup with
	 contributions `felonious'
	\item David DeRemer: Safeguards have dropped off in use around the world
		\begin{itemize}
			\item Look at U.S. vs. ROW (he sent papers)
			\item Wouters and Zissimos (2016 WP): Appellate Body failed to give clear direction in cases in early 2000s as to when Safeguards can be applied. See Sykes (2003).
		\end{itemize}
	\item Limao and Tovar 2011
		\begin{enumerate}
			\item Their 2nd (and final) stage is Nash bargaining. Is there an isomorphism to my setup?
			\item It IS clear to take their first, limited case of commitment to tariff cap only (ignore NTB) as analogy for gov't in my model choosing commitment to different tariff level. Need to answer Q1 above in order to answer points 3 and 4:
			\item Their gains from commitment derive from improvement in bargaining power vis-a-vis the lobby (Schelling conjecture invoked). How does this relate to the gain from commitment in my model?
			\item Top of page 6: ``large enough bargaining power.'' Can I make connection to $\ga(e)$?
			\item Note: in their equation 3, $C$ enters expression for NBS's ($\tau$) in equilibrium, it doesn't disappear as in GH
		\end{enumerate}
\end{itemize}

@@
